Natural Stories model analyses

### 2022-08-15 Analysis/models.Rmd

```{r meaned data, eval=T, cache=T}
gam_ngram_1 <- gam(formula_interact, data=ngram_mean_data, method="REML")
gam_grnn_1 <- gam(formula_interact, data=grnn_mean_data, method="REML")
gam_txl_1 <- gam(formula_interact, data=txl_mean_data, method="REML")
gam_gpt_1 <- gam(formula_interact, data=gpt_mean_data, method="REML")
gam_gpt_1a <- gam(formula_interact, data=filter(gpt_mean_data,len<9), method="REML")
gam_gpt_2 <- gam(formula2_interact, data=gpt_mean_data, method="REML")
gam_gpt_3 <- gam(formula3_interact, data=gpt_mean_data, method="REML")

gam_gpt_3_hf <- gam(formula3_interact, data=subset(gpt_mean_data,freq > 2), method="REML")
gam_gpt_3_lf <- gam(formula3_interact, data=subset(gpt_mean_data,freq <= 2), method="REML")

gam_gpt_2_hf <- gam(formula2_interact, data=subset(gpt_mean_data,freq > 2), method="REML")
gam_gpt_2_lf <- gam(formula2_interact, data=subset(gpt_mean_data,freq <= 2), method="REML")

@@

summary(gam_gpt_no_interact_plus_linear_surprisal)
```

>>>>

```{r meaned data, eval=T, cache=T}
gam_ngram_1 <- gam(formula_interact, data=ngram_mean_data, method="REML")
gam_grnn_1 <- gam(formula_interact, data=grnn_mean_data, method="REML")
gam_txl_1 <- gam(formula_interact, data=txl_mean_data, method="REML")
gam_gpt_1 <- gam(formula_interact, data=gpt_mean_data, method="REML")
#gam_gpt_1a <- gam(formula_interact, data=filter(gpt_mean_data,len<9), method="REML")
gam_gpt_2 <- gam(formula2_interact, data=gpt_mean_data, method="REML")
gam_gpt_3 <- gam(formula3_interact, data=gpt_mean_data, method="REML")

#gam_gpt_3_hf <- gam(formula3_interact, data=subset(gpt_mean_data,freq > 2), method="REML")
#gam_gpt_3_lf <- gam(formula3_interact, data=subset(gpt_mean_data,freq <= 2), method="REML")

#gam_gpt_2_hf <- gam(formula2_interact, data=subset(gpt_mean_data,freq > 2), method="REML")
#gam_gpt_2_lf <- gam(formula2_interact, data=subset(gpt_mean_data,freq <= 2), method="REML")

@@

summary(gam_gpt_no_interact_plus_linear_surprisal)
```
## Extract linearity comparisons

```{r}
make_gam_linear_summary <- function(data){
  model <- gam(formula_no_interact_plus_linear_surprisal, data=data, method="REML")
  summ <- summary(model)
  stable <- summ$s.table %>% as_tibble(rownames="Term") %>% rename(pvalue=`p-value`) %>% select(Term, pvalue)
  ptable <- summ$p.table %>% as_tibble(rownames="Term") %>% rename(pvalue=`Pr(>|t|)`) %>% select(Term, pvalue)
  return(union(stable,ptable))
}

gpt_linear <- make_gam_linear_summary(gpt_mean_data) %>% mutate(model="GPT-2")
ngram_linear <- make_gam_linear_summary(ngram_mean_data) %>% mutate(model="5-gram")
txl_linear <-  make_gam_linear_summary(txl_mean_data) %>% mutate(model="TXL")
grnn_linear <-  make_gam_linear_summary(grnn_mean_data) %>% mutate(model="GRNN")

all_linear <- gpt_linear %>% union(ngram_linear) %>% union(txl_linear) %>% union(grnn_linear) %>% mutate(pvalue=pvalue(pvalue,accuracy=.0001,add_p=T),model=factor(model, levels=c("5-gram", "GRNN", "TXL","GPT-2")))%>% arrange(model) %>% 
  pivot_wider(names_from=model, values_from=pvalue) %>% write_rds(here("Analysis/models/gams_linearity_test.rds"))
```

## Pretty interaction pics
```{r fig.width=6, fig.height=8,}

#gam_gpt_3 <- gam(formula3_interact, data=gpt_mean_data, method="REML")

#gam_gpt_interact <- getViz(gam_gpt_3) %>% write_rds(here("Analysis/models/gam_gpt_interact.rds"))

# can't figure out how to adjust the number of rows cols using the default, so we instead copy their internal functions off github and hack

addDefaultLayers <- function( .l ){
    .cl <- paste(.l$type, collapse = '')

    if( grepl("Check", .cl) ){ # [A] Checking plots
      if("qgam" %in% mcls){ # a.1 Quantile GAMS
        .l <- switch(.cl, 
                     "Check0DScalarNumeric" = .l + l_hist() + l_vline(),
                     "Check0DScalarFactor" = .l + l_hist() + l_vline(),
                     "Check0DScalarLogical" = .l + l_hist() + l_vline(),
                     "Check0DVectorNumeric" = .l + l_hist(),
                     "Check0DVectorFactor" = .l + l_hist(),
                     "Check0DVectorLogical" = .l + l_hist(),
                     "Check1DNumeric" =  .l + l_gridQCheck1D() + l_rug(), 
                     "Check1DFactor" = .l + l_gridQCheck1D() + l_rug(),
                     "Check1DLogical" = .l + l_gridQCheck1D() + l_rug(),
                     "Check2DNumericNumeric" = .l + l_gridQCheck2D(),
                     "Check2DFactorFactor" = .l + l_gridQCheck2D() + l_rug(),
                     "Check2DFactorNumeric" = .l + l_gridQCheck2D() + l_rug(),
                     .l
        )
      } else { # a.2 Standard GAMS
        .l <- switch(.cl, 
                     "Check0DScalarNumeric" = .l + l_hist() + l_vline(),
                     "Check0DScalarFactor" = .l + l_hist() + l_vline(),
                     "Check0DScalarLogical" = .l + l_hist() + l_vline(),
                     "Check0DVectorNumeric" = .l + l_hist(),
                     "Check0DVectorFactor" = .l + l_hist(),
                     "Check0DVectorLogical" = .l + l_hist(),
                     "Check1DNumeric" =  .l + l_dens2D("cond") + l_gridCheck1D(mean, showReps = FALSE), 
                     "Check1DFactor" = .l + l_gridCheck1D(showReps = FALSE) + l_rug(),
                     "Check1DLogical" = .l + l_gridCheck1D(showReps = FALSE) + l_rug(),
                     "Check2DNumericNumeric" = .l + l_gridCheck2D(),
                     "Check2DFactorFactor" = .l + l_gridCheck2D() + l_rug(),
                     "Check2DFactorNumeric" = .l + l_gridCheck2D() + l_rug(),
                     .l
        )
      }
    } else { # [B] Smooth effect plots

    .l <- switch(.cl, 
                 "singleIndex1D" = .l + l_fitLine() + l_ciLine() + l_rug(), 
                 "fs1D" = .l + l_fitLine() + theme(legend.position="none"),
                 "1D" = .l + l_fitLine() + l_ciLine() + l_rug(),
                 "2D" = .l + l_fitRaster() + l_fitContour(), 
                 "MD" = .l + l_fitRaster() + l_fitContour(), 
                 "MDslice" = .l + l_fitRaster() + l_fitContour(), 
                 "sos0" = .l + l_fitRaster() + l_fitContour(), 
                 "sos1" = .l + l_fitRaster() + l_fitContour(),
                 "randomEffect" = .l + l_fitLine() + l_ciLine() + l_points(), 
                 "MultiRandomEffect" = .l + l_points(), 
                 "mrf" = .l + l_poly(), 
                 "PtermNumeric" = .l + l_fitLine() + l_ciLine() + l_rug(),
                 "PtermFactor" = .l + l_ciBar() + l_fitPoints() + l_rug(),
                 "PtermLogical" = .l + l_ciBar() + l_fitPoints() + l_rug(), 
                 "PtermMatrixNumeric" = .l + l_fitLine() + l_ciLine(),
                 "Multi1D" = .l + l_fitLine(),
                 "Multi2D" = .l + l_fitRaster() + l_fitContour(),
                 "MultiPtermNumeric" = .l + l_ciBar() + l_fitPoints(),
                 "MultiPtermFactor" = .l + l_ciBar() + l_fitPoints(), 
                 "ALE1DNumeric" = .l + l_fitLine() + l_ciLine() + l_rug(),
                 "ALE1DFactor" = .l + l_ciBar() + l_fitPoints() + l_rug(),
                  .l
                 )

    return( .l )
  }

g <- read_rds(here("Analysis/models/gam_gpt_interact.rds")) %>% plot()
g$plots <- lapply(g$plots,   function(.l){
                         addDefaultLayers(.l)
                       })

blah <-  arrangeGrob(grobs=read_rds(here("Analysis/models/draft_grid.rds")), ncol=3, )

ggsave(here("Analysis/models/gam_grid.png"), blah)
